---
title: "Emergency Room - Annie Nguyen"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Patients arrive at the Emergency Room following an unknown probability distribution
(stream 1). They will be treated by either of two doctors.

A proportion of the patients are classified as NIA (need immediate attention) and the
rest as CW (can wait). NIA patients are given the highest priority, 3, see a doctor as soon
as possible for 40 ± 30 minutes (stream 2), then have their priority reduced to 2 and wait
until a doctor is free again, when they receive further treatment for 30 ± 20 minutes (stream
3) and are discharged.

CW patients initially receive a priority of 1 and are treated (when their turn comes) for
15 ± 10 minutes (stream 4); their priority is then increased to 2, they wait again until a
doctor is free, receive 10 ± 5 minutes (stream 5) of final treatment and are discharged.

An important aspect of this system is that patients who have already seen the doctor
once compete with newly arriving patients who need a doctor. As indicated, patients who
have already seen the doctor once have a priority level of 2 (either increased from 1 to 2 or
decreased from 3 to 2). Thus, there is one shared queue for the first treatment activity and
the final treatment activity. In addition, we assume that the doctors are interchangeable.
That is, it does not matter which of the two doctors performs the first or final treatment.
Simulate for 20 days of continuous operation, 24 hours per day. Note, the inter-arrival
time and type of 100 patients are collected (based on a randomly selected weekday data).
The data is attached.

1. Analyze your results and explain your suggestions for reducing the waiting time of the
patients.

2. What is the average flow-time for NIA and CW patients before or after applying
suggestions. different suggestions.

3. Discuss the utilization of doctors before or after applying suggestions.

```{r cars}
library(simmer)


#Define trajectory for patient
patient <- trajectory("patients' path") %>%
  
  branch(option = function() sample(1:2, size=1, prob=c(0.18, 0.82), replace=T), continue = c(T,T), #18% of patients are NIA and 82% are CW (analyzed sample dataset). Define 2 branches.
         
  trajectory("NIA") %>% #Sub-trajectory for NIA patients
    
    set_attribute("priority", 3) %>% #Set to priority 3
    set_prioritization(c(3, 7, T)) %>%
    seize("doctor", 1) %>%
    timeout(function() runif(1, 10, 70)) %>% #Uniform distribution 40 ± 30 minutes (between 10 and 70 minutes)
    release("doctor", 1) %>%
    
    set_attribute("priority", 2) %>% #Decrease to priority 2
    set_prioritization(c(2, 7, T)) %>%
    seize("doctor", 1) %>%
    timeout(function() runif(1, 10, 50)) %>% #Uniform distribution 30 ± 20 minutes (between 10 and 50 minutes)
    release("doctor", 1),
  
  trajectory("CW") %>% #Sub-trajectory for CW patients
    
    set_attribute("priority", 1) %>% #Set to priority 1
    set_prioritization(c(1, 7, T)) %>%
    seize("doctor", 1) %>%
    timeout(function() runif(1, 5, 25)) %>% #Uniform distribution 15 ± 10 minutes (between 5 and 25 minutes)
    release("doctor", 1) %>%
     
    set_attribute("priority", 2) %>% #Increase to priority 2
    set_prioritization(c(2, 7, T)) %>%
    seize("doctor", 1) %>%
    timeout(function() runif(1, 5, 15)) %>% #Uniform distribution 10 ± 15 minutes (between 5 and 15 minutes)
    release("doctor", 1)
) 


set.seed(12)
envs <- lapply(1:100, function(i) { #100 replications
  simmer("Clinic") %>%
    add_resource("doctor", 2) %>% #2 doctors
    add_generator("patient", patient, function() rnorm(1, 15.077, 14.359), mon=2) %>% #Interarrival time assuming normal distribution with mean=15.077, sd=14.359 (Calculated from sample dataset)
  run(28800) #20 days * 24 hours * 60 minutes = 28800 minutes
})

#Visualize
library(simmer.plot)
plot(patient, verbose=T)



#a) 
#We can expect to reduce the waiting time of patients when additional resources (doctors) are added.

#Waiting time plot before adding additional resource
arrivals <-get_mon_arrivals(envs)
plot(arrivals, metric="waiting_time")

#Average waiting time overall before adding additional resource
waitingTime <- (arrivals$end_time - arrivals$start_time) - arrivals$activity_time
mean(waitingTime)
#The average waiting time for both NIA and CW patients is 14.19 minutes.


#Add additional resource (+ one doctor)
set.seed(123)
envs2 <- lapply(1:100, function(i) { #100 replications
  simmer("Clinic") %>%
    add_resource("doctor", 3) %>% #3 doctors
    add_generator("patient", patient, function() rnorm(1, 15.077, 14.359), mon=2) %>% #Interarrival time assuming normal distribution with mean=15.077, sd=14.359 (Calculated from sample dataset)
  run(28800) #20 days * 24 hours * 60 minutes = 28800 minutes
})

#Waiting time plot after adding additional resource
arrivals2 <-get_mon_arrivals(envs2)
plot(arrivals2, metric="waiting_time")

#Average waiting time overall after adding additional resource
waitingTime2 <- (arrivals2$end_time - arrivals2$start_time) - arrivals2$activity_time
mean(waitingTime2)

#When we add an additional resource (+ one doctor), as expected, the waiting time for both NIA and CW patients is reduced to 1.68 minutes.



#b)
#Flow time plot before adding additional resource
plot(arrivals, metric="flow_time")

#Calculate flow time for each patient before adding additional resource
x1 <- get_mon_arrivals(envs) 
x2 <- get_mon_attributes(envs) 
x1
x2

all <- merge(x1, x2, by=c("name", "replication"), all = T)
all <- na.omit(all)

NIA <- subset(all, all$value == 1)
CW <- subset(all, all$value == 2)

NIA.flowTime = (NIA$end_time - NIA$start_time)
CW.flowTime = (CW$end_time - CW$start_time)

#Average flow time for each type of patient before adding additional resource
mean(NIA.flowTime, na.rm = T)
#Average flow time is 40.26 minutes for NIA patients.
mean(CW.flowTime, na.rm = T)
#Average flow time is 48.09 minutes for CW patients.


#Calculate flow time for each patient after adding additional resource
x3 <- get_mon_arrivals(envs2) 
x4 <- get_mon_attributes(envs2) 
x3
x4

all <- merge(x3, x4, by=c("name", "replication"), all = T)
all <- na.omit(all)

NIA2 <- subset(all, all$value == 1)
CW2 <- subset(all, all$value == 2)

NIA2.flowTime = (NIA2$end_time - NIA2$start_time)
CW2.flowTime = (CW2$end_time - CW2$start_time)

#Average flow time for each type of patient after adding additional resource
mean(NIA2.flowTime, na.rm = T)
#Average flow time is 26.37 minutes for NIA patients.
mean(CW2.flowTime, na.rm = T)
#Average flow time is 34.28 minutes for CW patients.

#When we add an additional resource (+ one doctor), the flow time is reduced significantly for both type of patients. 



#c)
#Utilization plot for 2 doctors (before)
resources <- get_mon_resources(envs)
head(resources)
plot(resources, metric="utilization") 

#Utilization plot for 3 doctors (after)
resources2 <- get_mon_resources(envs2)
head(resources2)
plot(resources2, metric="utilization") 

#When we add an additional resource (+ one doctor), the utilization percentage for our resource decreases.
#This is expected because the more resources there are, the less occupied they will be.
```
